---
title       : Project Presentation
subtitle    : Developing Data Products
author      : Mohamed Taufeeq
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
---
## Ideal Body Weight (IBW)

To determine how much you should weigh (your ideal body weight) several factors should be considered, including age, muscle-fat ratio, height, sex, and bone density. 

This app will help you to detemine your IBW using the three standard approaches.


Body Mass Index (BMI) 
=====================
Your BMI is your weight in relation to your height. 
```{r BMI, cache=TRUE, echo=TRUE}
weight <- 80; height <- 1.8
BMI <- weight / (height ^2)
if (BMI >= 18.5 && BMI <= 25) {
                print("Your weight is ideal")
        }   
```

---

Waist-Hip Ratio (WHR) 
=====================
A waist-hip measurement is the ratio of the circumference of your waist to that of your hips.
```{r WHR, cache=TRUE, echo=TRUE}
waist <- 28; hip <- 36
WHR <- waist / hip
WHR
```

Waist-to-Height Ratio
=====================
```{r, cache=TRUE, echo=TRUE}
waist <- 91
height <- 183
if (waist <= (height / 2)) {
                print("This can help you to increase your life expectancy.")
        }
        if (waist > (height / 2)) {
                print("Your WHtR is not ideal.")
        }
```

---

## ui.R chunk
Below is the code snippet of the ui.R
```{r UI, cache=TRUE, echo=TRUE}
library(shiny)

shinyUI(pageWithSidebar(
        
        headerPanel("Ideal body weight"),
        
        sidebarPanel(                
                h3('Body Mass Index(BMI)'),
                p('Your BMI is your weight in relation to your height.'),
                
                numericInput(
                        'weight',
                        'Weight in kilogram(kg)',
                        0,
                        min = NA,
                        max = NA,
                        step = NA
                ),
                
                numericInput(
                        'height',
                        'Height in centi-meter(cm)',
                        0,
                        min = NA,
                        max = NA,
                        step = NA),
                
                submitButton('Your BMI')
                ),
        mainPanel(
                h3('Result of BMI Measurement'),
                
                h4('Your weight in kg'),
                p('Example: 95'),
                verbatimTextOutput("weight")
                )
        ))
```

---

## server.R chunk

Here the code snippet of the server.R
```{r server, cache=TRUE, echo=TRUE}
shinyServer(
        function(input, output) {
                output$weight <- renderPrint({input$weight})
                output$height <- renderPrint({input$height})
                output$height1 <- renderPrint({input$height1})
                output$waist <- renderPrint({input$waist})
                output$waist1 <- renderPrint({input$waist1})
                output$hip <- renderPrint({input$hip})
                
                #call required function
                output$BMI <- renderPrint({BMI(input$weight, input$height)})   
                output$WHR <- renderPrint({WHR(input$waist, input$hip)})
                output$WHtR <- renderPrint({WHtR(input$waist1, input$height1)})
        }       
)
```